---
title: "SEISMIC WG1-P4"
date: "Updated: `r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    keep_md: true
    code_folding: hide
    theme: cosmo
    toc: yes
    toc_depth: 2
    toc_float: yes
subtitle: Analysis Workflow
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### - R/RStudio used for analyses
#### - Jupyter Notebooks/Google Co-Lab used for sharing code

# **I. Data Processing (Institution Specific)**

#### *Note:* 

- Exact syntax for these steps is likely to vary based on institution-specific variable naming conventions
- Sample code found on the [SEISMIC GitHub repository](https://github.com/seismic2020/) for complete examples of institution-specific data cleaning code

## 0. Startup

### a.  Load R pkgs
```{r message = FALSE}
# Finds and loads necessary libraries for the code to run
if (!require("pacman")) install.packages("pacman")
library(pacman)
pacman::p_load("tidyverse")   # Data wrangling

# etc...
```

### b.	Load full dataset
```{r include = FALSE}

# NOTE: Rmd headings
 ## {r } - runs and shows output and code
 ## {r echo = FALSE} - runs and shows output but doesnt show code
 ## {r include = FALSE} - runs, but doesnt show output or code 
 ## {r eval = FALSE} - doesn't run code (good for sample segments)
 ## {r message = FALSE, warning  = FALSE} - hides error and non-output messages (NOT RECOMMENDED!)

```

```{r eval = FALSE}
# CHANGE TO YOUR FILE PATH
df_full <- read.csv("~/YOUR FILE PATH HERE.csv")
head(names(df_full))
```

```{r echo = FALSE}
print("Example RAW Varnames")
head(names(df_full), 15)
```

## 1.	Clean Student level variables

### a. Rename and generate/recode student level variables as needed to match common SEISMIC AP variable names

#### *Note:* 

- Student level data should contain unique rows per student

```{r include = FALSE}
# CODE TO RUN GOES HERE
```

```{r eval = FALSE}
# SAMPLE CODE 
## Student Level ####
df_std <- df_full %>%
  # Renamed variables
  mutate(st_id = EMPLID_H) %>%
  mutate(ethniccode = ETHNIC_GROUP_CD) %>%
  mutate(famincome = abs(AGI)) %>%
  # Recoded variables
  mutate(firstgen = recode(FIRST_GENERATION_DESCR, "First Generation" = 1, "Not First Generation" = 0, "Unknown" = 0)) %>%
  mutate(ethniccode_cat = recode(ETHNIC_GROUP_CD, "HISPA" = 1, "BLACK" = 1, "AMIND" = 1, "PACIF" = 1, "ASIAN" = 2, "WHITE" = 0)) %>%
  mutate(urm = recode(ETHNIC_GROUP_CD, "HISPA" = 1, "BLACK" = 1, "AMIND" = 1, "PACIF" = 1, "ASIAN" = 0, "WHITE" = 0)) %>%
  mutate(gender = recode(GENDER_CD, "F"=1, "M"=0, "m"=0, "U" = 2)) %>%
  mutate(female = recode(GENDER_CD, "F"=1, "M"=0, "m"=0, "U" = 2)) %>%
  mutate(lowincomeflag = if_else(is.na(AGI), 0,
                                if_else(AGI <= 46435, 1,0))) 

# etc...
```

```{r echo = FALSE}
print("Example Student Level Dataframe")
sample_n(as_tibble(df_std), 10)

```

## 2.	Clean Course level variables

### a.	Rename and generate/recode course level variables as needed to match common SEISMIC AP variable names

#### *Note:* 

- Course level data will likely contain multiple rows for each course, per student

```{r include = FALSE}
# CODE TO RUN GOES HERE
```

```{r eval = FALSE}
# SAMPLE CODE
## Course Level ####
df_crs <- df_full %>%
  # Renamed variables
  mutate(st_id = EMPLID_H) %>%
  mutate(crs_sbj = SUBJECT_CD) %>%
  mutate(crs_catalog = CATALOG_NBR) %>%
  mutate(crs_name	= CLASS_TITLE) %>%
  mutate(crs_retake = REPEAT_CD) %>%
  mutate(crs_term	= TERM_CD) %>%
  # Recoded variables
  mutate(numgrade = GRADE_POINTS/UNITS_TAKEN) %>%
  mutate(numgrade_w = if_else(COURSE_GRADE_CD == "W", 1, 0)) %>%
  separate(as.character("TERM_CD"), c("crs_YEAR", "crs_SEMESTER"), 3, remove = FALSE) %>%
  separate(as.character("crs_YEAR"), c("crs_DEC", "crs_YEAR"), 1) %>% 
  mutate(crs_term_yr = crs_YEAR) %>%
  mutate(crs_term_sem = crs_SEMESTER) %>%
  mutate(summer_crs = if_else(endsWith(as.character(TERM_CD),"7"), 1, 0))

# etc...
```

```{r echo = FALSE}
print("Example Course Level Dataframe")
as_tibble(df_crs)
```

### b.	For each subject course (1 and 2), create dataframe of only first time taking that course

#### *Note:* 

- This step selects down to only a single row per course, per student

```{r include = FALSE}
# CODE TO RUN GOES HERE
```

```{r eval = FALSE}
# SAMPLE CODE
# By Course (Taking only First Attempt) ####
# Bio
df_crs_bio1 <- df_crs %>%
  filter(crs_sbj == "BIOSC" & (crs_catalog == "0150")) %>% # | crs_catalog == "0715")) %>%
  #Only first time taking course
  group_by(st_id, crs_catalog) %>% 
  arrange(crs_term, .by_group= TRUE) %>%
  mutate(crs_retake_num = row_number()) %>%
  filter(crs_retake_num == 1) 

# etc...
# repeat for Bio2, Chem1, Chem2, Phys1, Phys2
```

```{r echo = FALSE}
print("Example Course Level Dataframe - First Course Only")
as_tibble(df_crs_bio1)

```

## 3.	Clean AP Level variables

### a.  For each AP subject, rename and generate/recode course level variables as needed to match common SEISMIC AP variable names

#### *Note:*

- Taking highest (max) AP score recieved; single row per AP exam, per student

``` {r include = FALSE}
# CODE TO RUN GOES HERE
```

``` {r eval = FALSE}
# SAMPLE CODE GOES HERE
##### AP Level ####
# By course ####
# Bio
df_ap_bio <- df_full %>%
  mutate(st_id = EMPLID_H) %>%
  mutate(aptaker = ifelse(is.na(BY), 0, 1)) %>%
  mutate(eligible_to_skip = ifelse(BY >= 4 & !is.na(BY), 1, 0)) %>%
  mutate(eligible_to_skip_2 = ifelse(BY == 5 & !is.na(BY), 1, 0)) %>%
  mutate(tookcourse = ifelse(
    SUBJECT_CD == "BIOSC" & (CATALOG_NBR == "0150") & # | CATALOG_NBR == "0715") & 
      COURSE_GRADE_CD != "W", 1, 0)) %>%
  mutate(tookcourse_2 = ifelse(
    SUBJECT_CD == "BIOSC" & (CATALOG_NBR == "0160") & # | CATALOG_NBR == "0716") & 
      COURSE_GRADE_CD != "W", 1, 0)) %>%
  #mutate(apyear = ?) %>%
  mutate(apscore = as.character(BY)) %>%
  mutate(apscore_full = ifelse(is.na(BY), 0, BY)) %>%
  select(st_id, aptaker:apscore_full) %>%
  group_by(st_id) %>%
  summarize_at(vars(-group_cols()),max)

# etc...
# repeat for Chem, Phys
```
   
```{r echo = FALSE}
print("Example AP Level Dataframe - Highest Score by Exam")
as_tibble(df_ap_bio)

```

## 4. Create stacked dataset

### a.  Join previously generated dataframes (Student, Course1, Course2, and AP) for each course subject (BIO, CHEM, PHYS) 

#### *Note:*

- Include new variable: “discipline” as flag for each subject
    + BIO
    + CHEM
    + PHYS
    
```{r include = FALSE}
# CODE TO RUN GOES HERE
```

```{r eval = FALSE}
# SAMPLE CODE
# Bio (N=3090)
df_bio <- df_std %>%
  right_join(df_crs_bio2, by = "st_id") %>%
  full_join(df_crs_bio1, by = "st_id") %>%
  full_join(df_ap_bio, by = "st_id") %>%
  mutate(discipline = "BIO") %>%
  mutate(skipped_course = ifelse(tookcourse == 0 & tookcourse_2 == 1, 1, 0)) %>%
  select(discipline, st_id:hsgpa, crs_sbj.x:current_major.x, crs_sbj.y:current_major.y, 
         aptaker, apscore, apscore_full, eligible_to_skip, 
         tookcourse, tookcourse_2, skipped_course) 

# etc...
# repeat for Chem, Phys
```

```{r echo = FALSE}
print("Example Stacked BIO Dataframe")
as_tibble(df_bio)
```

### b. Stack complete dataframes for each course subject (BIO, CHEM, PHYS), including "discipline" indicator variable
```{r include = FALSE}
# CODE TO RUN GOES HERE
```

```{r eval = FALSE}
# SAMPLE CODE
# Stacked dataframe with Bio, Chem, Phys
df_clean <- rbind(df_bio, df_chem, df_phys)
df_clean <- df_clean %>%
  rename_at(vars(ends_with(".x")), 
            ~(str_replace(., ".x", "_2"))) %>%
  rename_at(vars(ends_with(".y")), 
            ~(str_replace(., ".y", "")))

rm(df_ap_bio, df_ap_chem, df_ap_phys, 
   df_bio, df_chem, df_phys, 
   df_crs_bio1, df_crs_bio2, df_crs_chem1, df_crs_chem2, df_crs_phys1, df_crs_phys2,
   df_crs, df_std)

# etc...
```

```{r eval = FALSE}
# SAMPLE CODE
# Stacked dataframe with Bio, Chem, Phys
df_clean <- rbind(df_bio, df_chem, df_phys)

# etc...
```

#### *Note:*

- Should end up with dataset structured like this [Example Dataset](https://docs.google.com/spreadsheets/d/1Sj5kaFNGUkBhRoOH3cIPm-97UEBZmcFkbKGjzBbKWc0/edit?usp=drive_open&ouid=118183464940790632947)

```{r echo = FALSE}
print("Example RE-CODED Varnames")
head(names(df_clean), 15)
```

# **II. Data Analysis (Same Across Institutions)**

#### *Note:* 

- Syntax for these steps should be the same for all institutions (once data cleaning steps above are followed)
- Sample code shown here; use [Shared Analysis file on SEISMIC Github Repository](https://github.com/seismic2020/) for complete analysis code

## 0. Startup

### a.  Load R pkgs
```{r message = FALSE}
if (!require("pacman")) install.packages("pacman")
library(pacman)
pacman::p_load("tidyverse",   # Data wrangling
               "epiDisplay")  # Display OR for logistic regressions

# etc...
```

### b.	Load clean dataset

```{r include = FALSE}
# CHANGE TO YOUR FILE PATH
df_clean <- read.csv("~/YOUR FILE PATH HERE.csv")
head(names(df_clean))
```

```{r echo = FALSE}
print("Example RE-CODED Varnames")
head(names(df_clean), 15)
```

## 1. Sample selection (by RQ, for each Course)

### a. Filter for student level inclusion/exclusion criteria

#### *Note:*

- Use [Shared Analysis Syntax](https://github.com/seismic2020/) for shared sample selection procedure

``` {r include = FALSE} 
# CODE TO RUN GOES HERE
```

``` {r eval = FALSE} 
df_clean <- df_clean %>%
  # Include
  filter(transfer == 0) %>%
  filter(tookcourse_2 == 1) %>%
  filter(cohort >= 2013 & cohort <= 2018) %>%
  # Exclude
  filter(international == 0)

# etc...
```

### b.	Create subset dataframes for each analysis sample

#### *Note:* 

- Use [Shared Analysis Syntax](https://github.com/seismic2020/WG1-P4/tree/master/Shared%20Analysis) or see [Sample Descriptions](https://docs.google.com/spreadsheets/d/1rN8W_iz1mr7lEzBGfdTZHa45wKOSLiSF8VEpChCPsmE/edit#gid=129222174) for full model specifications

```{r include = FALSE}
# CODE TO RUN GOES HERE
```        

```{r eval = FALSE}
# Bio
# Took 2nd course in sequence
df_bio2 <- df_clean %>%
  subset(discipline == "BIO") %>%
  subset(apyear >= 2013)

# etc...
# repeat for Chem, Phys
```

```{r eval = FALSE}   
# Took AP 
df_BYtakers <- df_clean %>%
  subset(discipline == "BIO") %>%
  subset(aptaker == 1)

# etc...
# repeat for Chem, Phys
```

```{r eval = FALSE}
# Skip eligible
df_BYeligible <- df_clean %>%
  subset(discipline == "BIO") %>%
  subset(eligible_to_skip == 1)

# etc...
# repeat for Chem, Phys
```

```{r eval = FALSE}
# Skip eligible, at each eligble score (NOTE: THESE WILL DIFFER BY INSTITUTION!)
df_BYeligible.4 <- df_bio2 %>%
  subset(apscore == 4)
df_BYeligible.5 <- df_bio2 %>%
  subset(apscore == 5)

# etc...
# Repeat for Chem, Phys
```

## 2.	Run Models (for each Course)

#### *Note:*

- Use [Shared Analysis Syntax](https://github.com/seismic2020/) for full model specifications

### RQ1: What student characteristics are associated with student participation and success in AP courses for students enrolled at the selected universities?

### *RQ1a: Who takes AP exams?*

```{r include = FALSE}
# CODE TO RUN GOES HERE
```

```{r eval = FALSE}
# Model 1a: Credits ####
#Bio
m1.a_bio <- glm(aptaker ~ factor(firstgen) + factor(lowincomeflag)  + factor(female) + factor(urm) +
                 scale(hsgpa) + scale(mathsr) + scale(englsr) + factor(cohort),
               binomial(link = "logit"), df_bio2)
logistic.display(m1.a_bio)

# etc...
# Repeat for Chem, Phys
```

```{r echo = FALSE}
print("Example: Model Parameters for RQ1a - Bio")
logistic.display(m1.a_bio)
```

### *REPEAT FOR OTHER RQs!*

